---
title: 'Milestone #4'
author: "Stephanie Cheng"
date: "3/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(janitor)
library(magrittr)
```

```{r cleaning and loading data, echo=FALSE}

# To clean and load the data, I used read.csv to read the csv files. I then used
# clean_names() from the janitor package to make the column names easier to work
# with. Then I went through and renamed all the column names with appropriate
# variable names since the original names were in the form of questions that the
# questionnaire initially used (which are long and tedious to work with). I did
# this for both the 2017 and 2019 data, trying my best to be consistent across
# the data since the questionnaries for both years were slightly different.

census_2019 <- read.csv("./raw-data/design_census_2019.csv") %>%
  clean_names() %>%
  rename(career_duration = i_ve_worked_in_design_for,
         job_type = i_work_in,
         industry_type = industries_i_work_with,
         skills = my_skills_include,
         currently_learning = i_am_currently_learning,
         employment_status = i_am_currently_a_an,
         job_level = i_m_a_an,
         job_challenges = my_job_is_challenging_because,
         job_pros = my_job_is_great_because,
         org_size = my_organization_has,
         department_size = my_department_has,
         job_duration = i_ve_worked_here_for,
         role_duration = i_ve_had_my_role_for,
         work_hours = i_typically_work,
         salary = i_make,
         benefits = my_benefits_include,
         advancement_opportunities = i_have,
         learning_status = i_am,
         design_satisfaction_level = i_am_1,
         job_stability = i_feel,
         side_gigs = on_the_side_i,
         learned_design_from = i_learned_design_from,
         degree_subject = i_got_my_degree_in,
         number_of_jobs = i_ve_had,
         up_to_date_method = i_stay_current_with_design_by,
         job_satisfaction_level = i_m_currently,
         zip_code = my_zip_code_is,
         age = my_age_is,
         gender = i_identify_as,
         ethnicity = i_am_2,
         issues_in_design = i_feel_the_most_critical_issues_challenges_currently_facing_design_are,
         biggest_future_impact = i_think_the_emerging_future_technologies_that_will_have_the_biggest_impact_on_design_are,
         most_critical_skills = i_think_the_most_critical_design_skills_for_the_future_will_be,
         priorities = my_main_priorities_are) %>%
  select(-i_work_in_1, -i_work_in_2, -i_work_in_3, -i_m_not_currently_working_because, -i_feel_the_most_critical_issues_challenges_currently_facing_design_are_1) %>%
  mutate(census_year = 2019)

census_2017 <- read.csv("./raw-data/design_census_2017.csv") %>%
  clean_names() %>%
  rename(career_duration = starts_with("x1_"),
         job_type = starts_with("x2_"),
         skills = starts_with("x3_"),
         employment_status = starts_with("x4_"),
         company_type = starts_with("x5_"),
         org_size = starts_with("x9_"),
         department_size = starts_with("x10_"),
         industry_type = starts_with("x11_"),
         job_duration = starts_with("x12_"),
         job_level = starts_with("x13_"),
         role_duration = starts_with("x16_"),
         work_hours = starts_with("x17_"),
         salary = starts_with("x18_"),
         benefits = starts_with("x19_"),
         advancement_opportunities = starts_with("x20_"),
         learning_status = starts_with("x21_"),
         job_satisfaction_level = starts_with("x22_"),
         job_stability = starts_with("x23_"),
         side_gigs = starts_with("x24"),
         cups_of_coffee = starts_with("x25"),
         highest_education = starts_with("x26_"),
         learned_design_from = starts_with("x27_"),
         degree_subject = starts_with("x28_"),
         number_of_jobs = starts_with("x29_"),
         number_of_job_shifts = starts_with("x30_"),
         up_to_date_method = starts_with("x31_"),
         areas_of_improvement = starts_with("x33_"),
         job_status = starts_with("x34_"),
         zip_code = starts_with("x36_"),
         age = starts_with("x37_"),
         gender = starts_with("x38_"),
         lgbtqia_status = starts_with("x39_"),
         ethnicity = starts_with("x40_"),
         issues_in_design = starts_with("x41_"),
         biggest_future_impact = starts_with("x42_"),
         most_critical_skills = starts_with("x43_")) %>%
  select(-starts_with("x6_"),
         -starts_with("x7_"),
         -starts_with("x8_"),
         -starts_with("x14_"),
         -starts_with("x32_"),
         -starts_with("x35_"))

# the us zipcode and us company data did not need much cleaning! I simply
# selected useful columns from the zipcode data.

us_zips <- read.csv("./raw-data/uszips.csv") %>%
  select(zip, state_name, lat, lng)

us_companies <- read.csv("./raw-data/us_companies.csv")
```

## Initial Data Analysis and Further Cleaning
For the Milestone this week, I wanted to take a closer dive into the data
through making preliminary bar graphs of the distribution of designers across
states (over the two different design census years) as well as distribution of
US companies across states. I first cleaned up the bar graphs for the design
census by normalizing by the total number of census respondents each year so I
had a percent of the designers in each state (rather than just the raw count).
This was key so that I could compare the data across both years. For the US
companies data, I just left the data as raw count because I didn't have a direct
comparison just yet. Then I went ahead and joined the data for the two years of
design census and created a joined bar graph that compares the percent
distribution of designers in different states in each year. I feel pretty good
about this preliminary analysis and am excited to wrangle the rest of this
dataset as there are a multitude of other variables I can investigate! (My
detailed processes for making this graphics are in commented in the code)


```{r initial analysis, echo=FALSE}

# 2019 data

# For the design census data, I first joined them with the us zipcode data so I
# could gather the states of each respondent. I made sure to omit the NAs using
# na.omit so I could cleanly analyze the location data. I grouped by state_name
# and counted the number of respondents so I could get a tally of the number of
# designers in each state.

census_2019_by_state <- census_2019 %>%
  left_join(us_zips, by = c("zip_code" = "zip")) %>%
  select(zip_code, state_name) %>%
  na.omit() %>%
  group_by(state_name) %>%
  count() %>%
  arrange(desc(n))

# I then used the summarize function with a new variable that takes the total
# number of designers across all states. This variable will be useful for
# normalizing the bar plot data so I can do a cross comparison over the two
# different years.

census_2019_total <- census_2019_by_state %>%
  summarize(total_sum = sum(census_2019_by_state$n)) %>%
  slice(1) %>%
  pull(total_sum)
  
# I made a bar plot that displays all the states and the corresponding percent
# of designers so I can get a sense of the distribution. To ensure the plot
# displayed correctly, I had to set stat as "identity" in geom_bar() so it's not
# the default stat count. I made sure to do coord_flip() because the state names
# are long and fit better on the y-axis.

census_2019_bar <- census_2019_by_state %>%
  ggplot(aes(x = state_name, y = n/census_2019_total)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Distribution of Designers by State in 2019",
       y = "Percent of Designers",
       x = "State")

# 2017 data

# I repeated the same procedure below for the 2017 census data as well.

census_2017_by_state <- census_2017 %>%
  left_join(us_zips, by = c("zip_code" = "zip")) %>%
  select(zip_code, state_name) %>%
  na.omit() %>%
  group_by(state_name) %>%
  count() %>%
  arrange(desc(n))

census_2017_total <- census_2017_by_state %>%
  summarize(total_sum = sum(census_2017_by_state$n)) %>%
  slice(1) %>%
  pull(total_sum)

census_2017_bar <- census_2017_by_state %>%
  ggplot(aes(x = state_name, y = n/census_2019_total)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Distribution of Designers by State in 2017",
       y = "Percent of Designers",
       x = "State")

# US company data

us_companies_by_state <- us_companies %>%
  left_join(us_zips, by = c("zip_code" = "zip")) %>%
  select(zip_code, state_name) %>%
  na.omit() %>%
  group_by(state_name) %>%
  count() %>%
  ggplot(aes(state_name, n)) +
  geom_bar(stat = "identity") +
  labs(title = "US Companies Distribution by State",
       x = "State",
       y = "Number of companies") +
  coord_flip()

census_2019_bar
census_2017_bar
us_companies_by_state
```

```{r comparison over two years, echo=FALSE}

# To make the comparison bar graphs I initially slided for the top 10 most
# designer populated states since displaying all the states isn't very
# informative at this moment. I then used the mutate function to add a year
# column and a percent column that takes the count and divides by the total
# number of respondents that year and multiplies by 100 to get percentage
# values.


top_10_states_2019 <- as_tibble(census_2019_by_state) %>%
  slice(1:10) %>%
  mutate(year = rep("2019"),
         percent = n/census_2019_total*100)

top_10_states_2017 <- as_tibble(census_2017_by_state) %>%
  slice(1:10) %>%
  mutate(year = rep("2017"),
         percent = n/census_2017_total*100)

# I used rbind to stack the datasets on top of each other.

joined_top_10_states <- rbind(top_10_states_2019, top_10_states_2017)

# With the stacked datasets, I created a bar plot making sure to fill by year. I
# had to change the position of the bars to "dodge" so they wouldn't be stacked
# on top of each other in order to be more informative about the differences
# across years. Again, I did coord_flip to make the state names easier to read.

joined_bar <- joined_top_10_states %>%
  group_by(year) %>%
  ggplot(aes(x = state_name, y = percent, fill = year)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "State",
       y = "Percent of Designers",
       title = "Distribution of Designers in Top 10 States") +
  coord_flip()
  
joined_bar
```

```{r average income and work hours by state, echo=FALSE}

normalize_2019 <- tibble(count = rep(top_10_states_2019$n, 8),
                         state_name = rep(top_10_states_2019$state_name, 8)) %>%
  arrange(state_name)

census_2019_salary_hours <- census_2019 %>%
  left_join(us_zips, by = c("zip_code" = "zip")) %>%
  select(zip_code, state_name, salary, work_hours) %>%
  na.omit() %>%
  filter(state_name %in% top_10_states_2019$state_name) %>%
  group_by(state_name) 

census_2019_salary_hours$salary <- factor(census_2019_salary_hours$salary, levels = c("Less than $25,000", "$25,000 to $34,999", "$35,000 to $49,999", "$50,000 to $74,999", "$75,000 to $99,999", "$100,000 to $149,999", "$150,000 to $199,999", "$200,000+"))

# normalization attempt
  # count(salary) %>%
  # as_tibble() %>%
  # add_row(state_name = "Ohio", salary = "$200,000+", n = 0) %>%  
  # mutate(pct = n/normalize_2019$count) %>%
  # arrange(state_name)
  
  
census_2019_salary_by_state_bar <- census_2019_salary_hours %>%
  group_by(state_name) %>%
  ggplot(aes(x = salary, fill = state_name)) +
  geom_bar() +
  labs(x = "Salary",
       y = "Number of designers",
       title = "Distribution of Salary of Designers by State") +
  coord_flip()

census_2019_salary_by_state_bar
```



